---
title: "homework_4"
author: "Aurna Mukherjee (am96848)"
date: "2025-02-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(tidyverse)
library(kableExtra)
library(mosaic)
```

# **Problem 1**

```{r echo=FALSE}
#ask what to use the 70 value for
n_trades <- 2021  # Total trades
flag_rate <- 0.024  # Baseline flagging probability under null hypothesis
observed_flags <- 70  # Observed flagged trades
n_simulations <- 100000  # Number of Monte Carlo simulations

sim_bank = do(100000)*nflip(n_trades, flag_rate)
sim_bank

sum(sim_bank >= 70)/100000 #p value

ggplot() + geom_histogram(aes(x = sim_bank$nflip))
```

The null hypothesis is that over the long run, securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders. The test statistic is the number of security trades that are flagged. Since the P-Value is returning close to 0.00186 consistently, which is pretty small, it seems likely that illegal trades might be happening.

# **Problem 2**

```{r echo=FALSE}
#how to use 1500 number
sim_health = do(100000)*nflip(n=50, prob=0.03)
sum(sim_health >= 8)/100000

ggplot() + geom_histogram(aes(x = sim_health$nflip), binwidth = 1)
```

The null hypothesis is that on average, restaurants in the city are cited for health code violations at the same 3% baseline rate. The test statistic is the number of restaurants that get health code violations out of the total number of inspections. The P-Value I got is 1e-04, which makes me think that the null hypothesis may not be true and that Gourmet Bites is getting significantly higher healthcare violations.

# **Problem 3**

```{r echo=FALSE}
total_jurors <- 240  # Total number of jurors selected
expected_proportions <- c(0.30, 0.25, 0.20, 0.15, 0.10)  # County proportions
observed_counts <- c(85, 56, 59, 27, 13)  # Observed jurors per group

tibble(observed = observed_counts, expected = expected_proportions*240)

chi2 = chi_squared_statistic(simulated_counts, total_jurors*expected_proportions)

num_simulations = 100000
chi2_sim = do(num_simulations)*{
  simulated_counts = rmultinom(1, total_jurors, expected_proportions)
  this_chi2 = chi_squared_statistic(simulated_counts, total_jurors*expected_proportions)
  c(chi2 = this_chi2) # return a vector with names and values
}

ggplot(chi2_sim) + 
  geom_histogram(aes(x=chi2))


# my crazy bag
my_chi2 = chi_squared_statistic(observed_counts, total_jurors*expected_proportions)
my_chi2

sum(chi2_sim >= my_chi2)/100000
```

I created a tibble of the observed jury counts for each group and the expected jury counts. My null hypothesis is that there is no systematic bias in jury selection, and I calculated a chi statistic of 12.42639. From there, I got a P-Value of 0.01452. This could suggest that there is evidence of systematic bias in jury selection, and this could be ran multiple times to see if the P-Value changes.

# **Problem 4**

```{r echo=FALSE}
#Part A
library(tidyverse)
letter_frequencies = read.csv("letter_frequencies.csv")
line = readLines("brown_sentences.txt")

calculate_chi_squared = function(line, freq_table) {
  
  # Ensure letter frequencies are normalized and sum to 1
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  # Remove non-letters and convert to uppercase
  line = gsub("[^A-Za-z] ", "", line)
  line = toupper(line)
  observed_counts = table(factor(strsplit(line, "")[[1]], levels = freq_table$Letter))
  
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  # Chi-squared statistic
  expected_counts[expected_counts == 0] <- 1e-10
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}
chi_sq_values = sapply(line, calculate_chi_squared, freq_table = letter_frequencies) %>%
  as_tibble()
  

```

```{r echo=FALSE}
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)


observed_chi_sq <- sapply(sentences, calculate_chi_squared, freq_table = letter_frequencies)

num_simulations <- 10000
p_values <- numeric(length(sentences))

# Loop through each sentence
for (i in seq_along(sentences)) {
  # Simulate chi-squared values
  chi2_sim <- replicate(num_simulations, {
    # 1. Simulate data based on EXPECTED frequencies (Crucial!)
    simulated_counts <- rmultinom(1, sum(nchar(sentences[i])), prob = letter_frequencies[, 2])
    simulated_freq <- data.frame(Letter = letter_frequencies$Letter, Probability = simulated_counts / sum(simulated_counts))

    # 2. Calculate chi-squared for the SIMULATED data
    if (simulated_freq$Probability[i] == 0) {
      p_values[i] <- 0
    }

    x <- calculate_chi_squared(sentences[i], simulated_freq)
    
  }) 

  # Calculate p-value
  p_values[i] <- sum(chi2_sim >= observed_chi_sq[i]) / num_simulations
}

# Create a data frame with results
p_values_df <- tibble(
  sentence_id = 1:length(sentences),
  chi_sq = observed_chi_sq,
  p_value = p_values
)

# Print the results
print(p_values_df)

```
